---
title: "Experiment 2"
author: "Alexander (Sasha) Pastukhov"
date: "19 5 2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(boot)
library(brms)
library(extrafont)
library(fs)
library(ggbeeswarm)
library(knitr)
library(kableExtra)
library(tidyverse)

loadfonts(device="win", quiet = TRUE)
```

```{r Tabula rasa}
rm(list = ls())
```


## Utility functions

Function that bootstraps a bias-corrected acceleated 95% confidence interval.

```{r bootstrap function}
bootstrap_mean_CIs <- function(df, target_var, R=2000){
  sample_mean <- function(data, index){
    mean(data[index])
  }
  
  boot_dist <- boot(data= df[[target_var]], statistic = sample_mean, R=R)
  boot_ci <- boot.ci(boot_dist, type="bca")$bca[4:5]

  tibble(Vmean = boot_dist$t0,
         Vmedian = median(boot_dist$t),
         Vlower = boot_ci[1], 
         Vupper = boot_ci[2])
}
```


Custom plotting theme
```{r Custom plotting theme}
theme_ba <- function(){
  theme(text=element_text(family="Arial"), 
        axis.text.x = element_text(size = 8, colour = 'black'), 
        axis.text.y = element_text(size = 8, colour = 'black'), 
        axis.title.x = element_text(size = 10), 
        axis.title.y = element_text(size = 10), 
        panel.grid.minor.x = element_blank(), 
        panel.grid.minor.y =  element_line(size = 0.24), 
        axis.ticks = element_line(size = 0.24))
}
```

Colors
```{r colors}
two_colors <- c("#66c2a5", "#fc8d62")
two_light_colors <- c("#b2e0d2", "#fdc6b0")


prime_probe_colors <- c("#E6444F", "#00457D")
prime_probe_light_colors <- c("#f2a1a7", "#7fa2be")
```


# Import and Preprocessing

Filename is coded as `Asynchronous Onset - <layout> - <subjectID> - <timestamp>.csv`, where

* `<layout>` is either _Parallel_ (spheres are presented to the left and to the right of fixation) or _Coaxial_ (spheres presented above and below the fixation).
* `<timestamp>` is in format YYYY-mm-dd-HH-MM-SS. Can be converted to an actual datetime using lubridate::ymd_hms() function.


The CSV-file format is the following:

* **Session**: string. Timestamp of the experimental session in format YYYY-mm-dd-HH-MM-SS. Can be converted to an actual datetime using lubridate::ymd_hms() function. Same as in filename.
* **Participant**: string. Unique participants ID, as in filename.
* **Block**: int. Block index.
* **Trial**: int. Trial index.
* **WhatIsReported": str. Either "probe" (only probe was presented) or "both" (both prime and probe were presented).
* **StereoStrength**: float. Strength of the stereoscopic depth disambiguation in degreees. 
* **BiasedSide**: string. Sphere that was disambiguated. For _parallel_ layout it is either _left_ or _right_. For the _coaxial_ it is either _above_ or _below_ the fixation.
* **RotationBias**: string. Biased direction of rotation, either _left_ or _right_.
* **ProbeDuration**: float. Duration of the probe.
* **OnsetDelay**: float. Onset delay before the start of the block in seconds.
* **ProbeSOA**: float. Stimulus onset asynchrony for the probe display, relative to the onset of the prime.
* **Percept**: string. Reported percept or event. Can be 
  * **left**: both spheres (co)rotated to the left
  * **right**: both spheres (co)rotated to the right
  * **up**/**down**: Sphere counterrotated. For _parallel_ layout _up_ means that the spheres rotated "into the screen" (the left sphere to the right, the right sphere to the left) and _down_ means that spheres rotated "out of the screen" (the left sphere rotated to the left, the right to the right). For the _coaxial_ layout both are renamed to **counterrotation** after the data is loaded in R.
  * **unclear**: unclear perception
* **RT**: float. Response time in seconds.


```{r Preprocessing}
results <- tibble(Layout= c("Parallel", "Coaxial")) %>%
  # listing all files in folders
  group_by(Layout) %>%
  do(tibble(filename= as.character(dir_ls("Data/Experiment 2", glob= stringr::str_c("*Onset*", .$Layout, "*.csv"))))) %>%
  
  # 1. loading files one-by-one and merging into the table
  group_by(Layout, filename) %>%
  do(read_csv(.$filename, col_types = cols(Session = col_character(),
                                           Participant = col_character(),
                                           Block = col_integer(),
                                           Trial = col_integer(),
                                           WhatIsReported = col_character(),
                                           StereoStrength = col_double(),
                                           BiasedSide = col_character(),
                                           RotationBias = col_character(),
                                           ProbeDuration = col_double(),
                                           OnsetDelay = col_double(),
                                           ProbeSOA = col_double(),
                                           Percept = col_character(),
                                           RT = col_double()))) %>%
  ungroup() %>%
  dplyr::select(-filename) %>%

  # 3. Renaming up/down to counter-rotation for "coaxial" layout
  rowwise() %>%
  mutate(Percept= ifelse(Layout == "Coaxial" && Percept %in% c("up", "down"), "counterrotation", Percept)) %>%
  
  # Rounding probe SOA
  mutate(OriginalProbeSOA = ProbeSOA,
         ProbeSOA = round(ProbeSOA, 2)) %>%

  # restoring order 
  ungroup() %>%
  arrange(Participant, Session, Block, Trial)

sort(unique(results$ProbeDuration))
```

## Parallel layout, _prime_ and _probe_ state relative to the biasing cues

```{r parallel layout}
parallel <- 
  results %>%
  ungroup() %>%
  
  # parallel layout, reports on both prime and probe
  filter(Layout == "Parallel", WhatIsReported == "both") %>%
  
  # figuring out perceptual states for individual spheres
  mutate(LeftSphere =  case_when(Percept == "up" ~ "right",
                                 Percept == "down" ~ "left",
                                 TRUE ~ Percept), 
         RightSphere =  case_when(Percept == "up" ~ "left",
                                  Percept == "down" ~ "right",
                                  TRUE ~ Percept)) %>%
    
  # biased (even if strength is 0) sphere is denoted as Prime, ambiguous as Probe
  mutate(Prime = ifelse(BiasedSide == "left", LeftSphere, RightSphere), 
         Probe = ifelse(BiasedSide == "left", RightSphere, LeftSphere)) %>%
  
  # figuring out direction opposite to the bias
  mutate(AntiBias = ifelse(RotationBias == "left", "right", "left"))


bias_effect_per_participant <-
  parallel %>%
  pivot_longer(cols = c("Prime", "Probe"), names_to = "Sphere", values_to = "State") %>%
  mutate(InBiasedDirection = State == RotationBias)
  
bias_effect <-
  bias_effect_per_participant %>%
  group_by(Participant, StereoStrength, ProbeDuration, Sphere) %>%
  summarise(Pbias = mean(InBiasedDirection)) %>%
  group_by(StereoStrength, ProbeDuration, Sphere) %>%
  do(bootstrap_mean_CIs(., "Pbias"))
```

```{r fitting models}
set.seed(4694561)
bias_effect_fits <-
  bias_effect_per_participant %>%
  group_by(Sphere) %>%
  nest()

bias_effect_fit_prototype <- brms::brm(InBiasedDirection ~ StereoStrength * ProbeDuration + ProbeSOA + (1|Participant),
                                       family = "bernoulli",
                                       data = bias_effect_fits$data[[1]],
                                       chains = 0)
  

bias_effect_fits <-
  bias_effect_fits %>%
  mutate(fit = purrr::map(data, ~update(bias_effect_fit_prototype, newdata = ., chains=4, cores=future::availableCores(), save_all_pars=TRUE)))

bias_effect_fits <-
  bias_effect_fits %>%
  mutate(fit = purrr::map(fit, ~brms::add_criterion(., "waic")))
```

```{r}
predict_using_median <- function(fit, data_for_predictions){
  terms <- 
  brms::fixef(fit) %>% 
  data.frame() %>%
  rownames_to_column("Term") %>%
  select(Term, Estimate) %>%
  tibble::deframe()

  data_for_predictions$Estimate <- boot::inv.logit(terms['Intercept'] + terms['StereoStrength']* data_for_predictions$StereoStrength + terms['ProbeDuration'] * data_for_predictions$ProbeDuration + 
    terms['ProbeSOA'] * data_for_predictions$ProbeSOA + terms['StereoStrength:ProbeDuration'] * data_for_predictions$StereoStrength * data_for_predictions$ProbeDuration)

  data_for_predictions
}

data_for_predictions <-
  expand.grid(StereoStrength = seq(0, 1.0, by=0.05),
              ProbeDuration = unique(bias_effect_per_participant$ProbeDuration)) %>%
  mutate(ProbeSOA = median(bias_effect_per_participant$ProbeSOA))

bias_effect_fits <-
  bias_effect_fits %>%
  mutate(predictions = purrr::map(fit, ~predict_using_median(., data_for_predictions)))
         
predictions <-
  bias_effect_fits %>%
  select(Sphere, predictions) %>%
  unnest(predictions)

save(bias_effect, predictions, bias_effect_fits, file="experiment2.RData")
```


```{r}
load("experiment2.RData")
ggplot(data= bias_effect, 
         aes(x = StereoStrength, 
             y = Vmedian,
             color = Sphere, fill=Sphere, group = Sphere, shape=Sphere)) + 
  geom_line(data = predictions, aes(y=Estimate), size=1) +
  geom_errorbar(aes(ymin = Vlower, ymax = Vupper), width= 0.05) +
  geom_point(size= 2.1) +
  theme_ba() +
  theme(legend.position = "none") +
  scale_y_continuous(name="P(sphere rotation = bias)", limits = c(0, 1), breaks = c(0, 0.25, 0.5, 0.75, 1), labels = c("0", "0.25", "0.5", "0.75", "1")) +
  scale_x_continuous(name="Disambiguation, stereo cues strength (deg)", 
                     breaks = sort(unique(bias_effect_per_participant$StereoStrength)),
                     labels = c("0", "0.25", "0.5", "1")) +
  scale_color_manual(values = prime_probe_colors) + 
  scale_shape_manual(values = c(21, 22)) +
  scale_fill_manual(values = prime_probe_light_colors) + 
  facet_grid(.~ProbeDuration, labeller = labeller(ProbeDuration = function(value){glue::glue("{value} s")}))


ggsave('experiment2.pdf', units="cm", device=cairo_pdf, width=12, height=8)
```

```{r detailed analysis for terms}
no_stereo_fit_prototype <- brms::brm(InBiasedDirection ~ ProbeDuration + StereoStrength : ProbeDuration + ProbeSOA + (1|Participant),
                                     family = "bernoulli",
                                     data = bias_effect_fits$data[[1]],
                                     chains = 0)

no_duration_fit_prototype <- brms::brm(InBiasedDirection ~ StereoStrength + StereoStrength : ProbeDuration + ProbeSOA + (1|Participant),
                                       family = "bernoulli",
                                       data = bias_effect_fits$data[[1]],
                                       chains = 0)

no_soa_fit_prototype <- brms::brm(InBiasedDirection ~ StereoStrength + ProbeDuration + StereoStrength : ProbeDuration + (1|Participant),
                                     family = "bernoulli",
                                     data = bias_effect_fits$data[[1]],
                                     chains = 0)
  
no_interaction_fit_prototype <- brms::brm(InBiasedDirection ~ StereoStrength + ProbeDuration + ProbeSOA + (1|Participant),
                                          family = "bernoulli",
                                          data = bias_effect_fits$data[[1]],
                                          chains = 0)

reduced_models <- list("Stereo Strength" = no_stereo_fit_prototype, 
                       "Probe Duration" = no_duration_fit_prototype,
                       "Probe SOA" = no_soa_fit_prototype,
                       "Strength:Duration" = no_interaction_fit_prototype)

model_summary <- function(fullmodel, modeldata){
  # fitting reduced models
  reduced_fits <- purrr::map(reduced_models, ~update(., newdata=modeldata, chains=4, cores=future::availableCores(), save_all_pars=TRUE))
  reduced_fits <- purrr::map(reduced_fits, ~brms::add_criterion(., "waic"))
  
  # comparison via WAIC
  dwaic <- purrr::map2(reduced_fits, names(reduced_models),
                      ~brms::loo_compare(fullmodel, .x, criterion="waic") %>%
                        data.frame() %>%
                        rownames_to_column("Model") %>%
                        mutate(elpd_diff = ifelse(Model != "fullmodel", -elpd_diff, elpd_diff)) %>%
                        filter(elpd_diff != 0) %>%
                        mutate(Term = .y)  %>%
                        select(Term, elpd_diff, se_diff))
  dwaic_df <- bind_rows(dwaic)

  # full model weight (importance of the term)  
  reduced_model_weight_df <- 
    purrr::map_df(reduced_fits, ~brms::model_weights(fullmodel, .)) %>%
    slice(1) %>%
    pivot_longer(cols = everything(), names_to = "Term", values_to = "Weight") %>%
    mutate(Weight = round(Weight, 2))

  # Bayes Factor
  reduced_bf <- 
    purrr::map_df(reduced_fits, ~brms::bayes_factor(fullmodel, .)$bf) %>% 
    pivot_longer(cols=everything(), names_to = "Term", values_to = "BF")
    
  # Betas
  term_effects <-
    fixef(fullmodel) %>% 
    data.frame() %>%
    rownames_to_column("Term") %>%
    filter(Term != "Intercept") %>%
    mutate_if(is.numeric, round, digits=3) %>%
    mutate(Term = names(reduced_models))
  
  left_join(term_effects, dwaic_df, by="Term") %>% 
    left_join(reduced_model_weight_df, by="Term") %>%
    left_join(reduced_bf, by="Term")
}
```


```{r model 1} 
set.seed(5945889)
modelSummary1 <- model_summary(bias_effect_fits$fit[[1]], bias_effect_fits$data[[1]])
modelSummary1 %>%
  mutate(elpd_diff = round(elpd_diff, 2),
         se_diff = round(se_diff, 2), 
         dWAIC = glue::glue("{elpd_diff}±{se_diff}"),
         BF = round(BF, 2),
         `95% CI` = glue::glue("{Q2.5}..{Q97.5}")) %>%
  select(Term, Estimate, `95% CI`, dWAIC, Weight, BF)

```

```{r model 2} 
set.seed(6395496)
modelSummary2 <- model_summary(bias_effect_fits$fit[[2]], bias_effect_fits$data[[2]])
modelSummary2 %>%
  mutate(elpd_diff = round(elpd_diff, 2),
         se_diff = round(se_diff, 2), 
         dWAIC = glue::glue("{elpd_diff}±{se_diff}"),
         BF = round(BF, 2),
         `95% CI` = glue::glue("{Q2.5}..{Q97.5}")) %>%
  select(Term, Estimate, `95% CI`, dWAIC, Weight, BF)
  
```

## Proportion of corotation in direction of bias for both layouts
```{r}
corotation_per_participant <-
  results %>%
  ungroup() %>%
  filter(WhatIsReported == "both") %>%
  mutate(InBiasedDirection = Percept == RotationBias)

corotation_with_bias <-
  corotation_per_participant %>%
  group_by(Participant, Layout, StereoStrength, ProbeDuration) %>%
  mutate(NWithBias = sum(InBiasedDirection),
         NTotal = n(),
         PWithBias = NWithBias / NTotal) %>%
  group_by(Layout, StereoStrength, ProbeDuration) %>%
  do(bootstrap_mean_CIs(., "PWithBias"))
```


```{r fitting coupling models}
set.seed(5614694)
corotation_fits <-
  corotation_per_participant %>%
  group_by(Layout) %>%
  nest()

bias_effect_fit_prototype <- brms::brm(InBiasedDirection ~ StereoStrength * ProbeDuration + ProbeSOA + (1|Participant),
                                       family = "bernoulli",
                                       data = corotation_fits$data[[1]],
                                       chains = 0)

corotation_fits <-
  corotation_fits %>%
  mutate(fit = purrr::map(data, ~update(bias_effect_fit_prototype,
                                        newdata = .,
                                        chains=4,
                                        cores=future::availableCores(),
                                        save_all_pars=TRUE)))

corotation_fits <-
  corotation_fits %>%
  mutate(fit = purrr::map(fit, ~brms::add_criterion(., "waic")))
```


```{r}
predict_using_median <- function(fit, data_for_predictions){
  terms <- 
  brms::fixef(fit) %>% 
  data.frame() %>%
  rownames_to_column("Term") %>%
  select(Term, Estimate) %>%
  tibble::deframe()

  data_for_predictions$Estimate <- boot::inv.logit(terms['Intercept'] + terms['StereoStrength']* data_for_predictions$StereoStrength + terms['ProbeDuration'] * data_for_predictions$ProbeDuration + 
    terms['ProbeSOA'] * data_for_predictions$ProbeSOA + terms['StereoStrength:ProbeDuration'] * data_for_predictions$StereoStrength * data_for_predictions$ProbeDuration)

  data_for_predictions
}

data_for_predictions <-
  expand.grid(StereoStrength = seq(0, 1.0, by=0.05),
              ProbeDuration = unique(corotation_per_participant$ProbeDuration)) %>%
  mutate(ProbeSOA = median(corotation_per_participant$ProbeSOA))

corotation_fits <-
  corotation_fits %>%
  mutate(predictions = purrr::map(fit, ~predict_using_median(., data_for_predictions)))
         
corotation_predictions <-
  corotation_fits %>%
  select(Layout, predictions) %>%
  unnest(predictions)

save(corotation_with_bias, corotation_fits, corotation_predictions, file="experiment2-corotation.RData")
```

```{r}
ggplot(data= corotation_with_bias, 
         aes(x = StereoStrength, 
             y = Vmedian,
             color = Layout, fill=Layout, group = Layout, shape=Layout)) + 
  geom_line(data = corotation_predictions, aes(y=Estimate)) +
  geom_errorbar(aes(ymin = Vlower, ymax = Vupper), width= 0.05) +
  geom_point(size= 2.1) +
  theme_ba() +
  theme(legend.position = c(0.2, 0.8)) +
  scale_y_continuous(name="P(sphere rotation = bias)", limits = c(0, 1), breaks = c(0, 0.25, 0.5, 0.75, 1), labels = c("0", "0.25", "0.5", "0.75", "1")) +
  scale_x_continuous(name="Disambiguation, stereo cues strength (deg)", 
                     breaks = c(0, 0.25, 0.5, 1),
                     labels = c("0", "0.25", "0.5", "1")) +
  scale_shape_manual(values = c(21, 22)) +
  facet_grid(.~ProbeDuration, labeller = labeller(ProbeDuration = function(value){glue::glue("{value} s")}))


ggsave('experiment2-corotation.pdf', units="cm", device=cairo_pdf, width=12, height=8)
```

### Corotation tables


```{r corotation model 1} 
set.seed(8895945)
modelSummary1 <- model_summary(corotation_fits$fit[[1]], corotation_fits$data[[1]])
modelSummary1 %>%
  mutate(elpd_diff = round(elpd_diff, 2),
         se_diff = round(se_diff, 2), 
         dWAIC = glue::glue("{elpd_diff}±{se_diff}"),
         BF = round(BF, 2),
         `95% CI` = glue::glue("{Q2.5}..{Q97.5}")) %>%
  select(Term, Estimate, `95% CI`, dWAIC, Weight, BF)
```


```{r corotation model 2}
load("experiment2-corotation.RData")
set.seed(9594588)
modelSummary2 <- model_summary(corotation_fits$fit[[2]], corotation_fits$data[[2]])
modelSummary2 %>%
  mutate(elpd_diff = round(elpd_diff, 2),
         se_diff = round(se_diff, 2), 
         dWAIC = glue::glue("{elpd_diff}±{se_diff}"),
         BF = round(BF, 2),
         `95% CI` = glue::glue("{Q2.5}..{Q97.5}")) %>%
  select(Term, Estimate, `95% CI`, dWAIC, Weight, BF)
```
